Informing IPCC accounting of forest carbon using the global forest carbon database (ForC v4.0)

### 2023-05-15 doc/manuscript/ESSDmanuscript.Rmd
minor clean-up
#39

Aboveground biomass, which is typically <200 t C ha^-1^ but can exceed 700 t C ha^-1^ [@anderson-teixeira_carbon_2021], is defined by the IPCC as "all biomass of living vegetation above the soil including stems, stumps, branches, bark, seeds, and foliage" [@ipcc_good_2003;@ipcc_2006_2006].
IPCC's guidance is that the understory may be excluded the understory if it constitutes a "minor" component, *where quantitative definitions of "understory" and "minor" are not provided*<!--verify (issue 40)-->, but where a commonly applied minimum size sampling threshold for mature forests would be 10 cm stem diameter at breast height (DBH).
A recent study characterizing the contributions of trees in different DBH classes to ecosystem C stocks and fluxes found that trees 1 - 10 cm DBH contributed up to ~8% aboveground biomass, ~17% aboveground woody net primary productivity ($ANPP_{woody.stem}$), and ~20% woody mortality ($M_{woody}$) of mature closed-canopy forests worldwide [@piponiot_distribution_2022].
In regrowth forests, woodlands, or savannas, small trees and shrubs contribute a much larger proportion of C stocks and fluxes [@lutz_global_2018; @piponiot_distribution_2022; @refs_from_ForC], and, correspondingly, biomass estimates for these ecosystems tend include smaller size classes [e.g., @refs_from_ForC]. 
While IPCC guidance specifies that all living vegetation should be included in biomass estimates, forest censuses and biomass estimates do not consistently include life forms other than dicot trees (e.g., lianas, ferns, palms, bamboo), although thhese do tend to be censused when they consitute a large proportion of the biomass [@refs_from_ForC].
Further, it is important to note that the IPCC definition of aboveground biomass excludes standing dead wood, which is included in remote sensing biomass estimates [@duncanson_aboveground_2021].

>>>>

Aboveground biomass, which is typically <200 t C ha^-1^ but can exceed 700 t C ha^-1^ [@anderson-teixeira_carbon_2021], is defined by the IPCC as "all biomass of living vegetation above the soil including stems, stumps, branches, bark, seeds, and foliage" [@ipcc_good_2003;@ipcc_2006_2006].
IPCC's guidance is that the understory may be excluded the understory if it constitutes a "minor" component, *where quantitative definitions of "understory" and "minor" are not provided*<!--verify (issue 40)-->, but where a commonly applied minimum size sampling threshold for mature forests would be 10 cm stem diameter at breast height (DBH).
A recent study characterizing the contributions of trees in different DBH classes to ecosystem C stocks and fluxes found that trees 1 - 10 cm DBH contributed up to ~8% aboveground biomass, ~17% aboveground woody net primary productivity ($ANPP_{woody.stem}$), and ~20% woody mortality ($M_{woody}$) of mature closed-canopy forests worldwide [@piponiot_distribution_2022].
In regrowth forests, woodlands, or savannas, small trees and shrubs contribute a much larger proportion of C stocks and fluxes [@lutz_global_2018; @piponiot_distribution_2022; @hughes_biomass_1999], and, correspondingly, biomass estimates for these ecosystems tend include smaller size classes. 
While IPCC guidance specifies that all living vegetation should be included in biomass estimates, forest censuses and biomass estimates do not consistently include life forms other than dicot trees (e.g., lianas, ferns, palms, bamboo), although these do tend to be censused when they constitute a large proportion of the biomass [@refs_from_ForC].
Further, it is important to note that the IPCC definition of aboveground biomass excludes standing dead wood, which is included in remote sensing biomass estimates [@duncanson_aboveground_2021].
